brados: declarative, programmable object storage

### 2016-05-25 paper.tex
asdf
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\subsection{Motivating Application: CORFU}

@@

But beyond the convience of writing less code, it is far easier for the
programmer writing an interface such as CORFU to convinence herself of the
correctness of the high-level details of the implementation without being
distracted by issues related to physical design or the many other gotchas that
one must deal with when writing low-level systems software.

\subsection{Optimization}

Choosing implementation details
  - the refcount example is good. small data kept in ceph object bytestream
until it becomes large then use omap.

  - could use the dataflow graph to do interesting things like understand how
  to do prefecting of things like keys or data that are often only known at
  execution time

@@

As we have seen the subset of Bloom used in Brados is powerful enough to
express the semantics of the CORFU storage interface. We briefly sketch the
implementation of an additional interface found in Ceph that used by the
S3-compatible RADOS-Gateway service to perform object reference counting.

@@

The reference implementation in Ceph is 300 LOC written in C++. The version
written in Brados is shown in Listing~\ref{lst:rc}. Line 10 shows the
implementation of requesting a reference which is an idempotent merge of the
tag into the persistent \texttt{refs} collection.  Unlike the CORFU interface
that took highlighted how collections are unified with persistent storage, the
reference counting interface requires access to RADOS runtime facilities that
allow objects to be removed. This is modeled in Brados using the Bloom
\emph{channel} abstraction that is used to transmit facts between entities in
the system; we effectively send a message to the runtime with our removal
request. The \texttt{remove\_chn} channel may be populated with messages and
the RADOS runtime will processes these asynchronously.

@@

\section{Evaluation}

@@

{\bf Active storage.}
There is a wide variety of work related \emph{active storage} that seeks to
increase performance and reduce data movement by exploiting remote storage
resources. This concept has been applied in the context of on-disk
controllers and object-based storage
(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
concerns have examined using managed lanuages, sandbox technologies, as well
as restricting extension installation to
vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.

@@

Much work remains to be done.  While we demonstrated some of the immediate
benefits of separating logical and physical specifications of component
compositions, the ``dream’’ of data independence is to take the physical
design question completely off the table for programmers.  We are currently
exploring both cost-based optimizers and autotuners to automate the selection
of a high performance physical design consistent with a given logical
specification.

The database literature on materialized view maintenance and selection can be
brought to bear in the context of programmable storage as well.  Recall the
sequencer logic of Corfu described in Section XX.  The Bloom specification
(Figure YY) simply states that the current value of the sequencer is the
result of evaluating an aggregate query over the log relation.  This means
that the system may choose any strategy to maintain a sequence value that is
consistent with this constraint.  The physical design decision to store and
update a memory-resident sequence value rather re-computing the maximal value
at each request is a view materialization decision that could similarly be
made by a compile-time optimizer.

>>>>

\subsection{Motivating Application: CORFU}
\label{sec:corfu}

@@

But beyond the convience of writing less code, it is far easier for the
programmer writing an interface such as CORFU to convenience herself of the
correctness of the high-level details of the implementation without being
distracted by issues related to physical design or the many other gotchas that
one must deal with when writing low-level systems software.

@@

As we have seen the subset of Bloom used in Brados is powerful enough to
express the semantics of the CORFU storage interface. We briefly sketch the
implementation of an additional interface found in Ceph that is used by the
S3-compatible RADOS-Gateway service to perform object reference counting.

@@

The reference implementation in Ceph is 300 LOC written in C++. The version
written in Brados is shown in Listing~\ref{lst:rc}. Line 10 shows the
implementation of requesting a reference which is an idempotent merge of the
tag into the persistent \texttt{refs} collection.  Unlike the CORFU interface
that highlighted how collections are unified with persistent storage, the
reference counting interface requires access to RADOS runtime facilities that
allow objects to be removed. This is modeled in Brados using the Bloom
\emph{channel} abstraction that is used to transmit facts between entities in
the system; we effectively send a message to the runtime with our removal
request. The \texttt{remove\_chn} channel may be populated with messages and
the RADOS runtime will processes these asynchronously.

@@

\subsection{Optimizations}

The challenge of navigating the physical design space has served as the
primary source of motivation for selection of a declarative langauge. While
our implementation does not yet

The Bloom language that we use as a basis for Brados produces a data flow
graph that can be used in static analysis. We envision that this graph will be
made available to the OSD and used to reorder and collesce requests based on
optimization criteria available from a performance model combined with
semantic information from the dataflow. For example today object classes are
represented as black boxes from the point of view of the OSD execution engine.
Understanding the behavior of an object class may allow intelligent
prefetching. Another type of analysis that may be useful for optimization is
optimistic execution combined with branch prediction where frequent paths
through a dataflow are handled optimistically.

@@

{\bf Active storage.}
There is a wide variety of work related to \emph{active storage} that seeks to
increase performance and reduce data movement by exploiting remote storage
resources. This concept has been applied in the context of on-disk
controllers and object-based storage
(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
concerns have examined using managed lanuages, sandbox technologies, as well
as restricting extension installation to
vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.

@@

Much work remains to be done.  While we demonstrated some of the immediate
benefits of separating logical and physical specifications of component
compositions, the ``dream'' of data independence is to take the physical
design question completely off the table for programmers.  We are currently
exploring both cost-based optimizers and autotuners to automate the selection
of a high performance physical design consistent with a given logical
specification.

The database literature on materialized view maintenance and selection can be
brought to bear in the context of programmable storage as well.  Recall the
sequencer logic of Corfu described in Section~\ref{sec:corfu}.  The Bloom specification
(Listing~\ref{lst:write}) simply states that the current value of the sequencer is the
result of evaluating an aggregate query over the log relation.  This means
that the system may choose any strategy to maintain a sequence value that is
consistent with this constraint.  The physical design decision to store and
update a memory-resident sequence value rather re-computing the maximal value
at each request is a view materialization decision that could similarly be
made by a compile-time optimizer.
